---
title: "lab_02"
author: "Derek Willis, adapted by Aidan Hughes"
date: "2024-08-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
library(lubridate)
```

## Load Data

For this lab we'll be working with a slice of Maryland voter registration data: records of people who have registration dates this year through Aug. 7. Let's load it.

```{r}
md_voters_2024 <- read_rds("data/lab_02.rds")
```

## Examine Data

Take a look at the data, getting familiar with what it contains. Most of the columns should be self-explanatory, but there are two date columns: STATE_REGISTRATION_DATE and COUNTY_REGISTRATION_DATE. The former represents the date the person registered as a Maryland voter; the latter is when they registered in their current county. They may be the same, but if a voter moves from one county to another they will differ. This is important because if the two dates are the same, that's a pretty good indication that this is a *new* voter, someone who has no history of registering in Maryland. You'll need to keep that definition in mind.

## Answer questions

**Q1** How many voters appear to be *new* registrants since July 21, when President Joe Biden announced he would leave the race, based on the information above? Create a dataframe to hold the only those voters.

Which counties have seen the most? Take a look at county population figures here and compare that list with your data findings in your answer: https://worldpopulationreview.com/us-counties/maryland.

**A1** 17076

```{r}
new_voters <- md_voters_2024 |>
filter(STATE_REGISTRATION_DATE == COUNTY_REGISTRATION_DATE) |>
filter(COUNTY_REGISTRATION_DATE > mdy("07/21/2024")

nrow(new_voters)

new_voters |>
group_by(COUNTY) |>
summarize(total_new = n()) |>
arrange(desc(total_new))
```

**Q2** A better question is whether the new voters you found in Q1 are an outlier compared to other weeks this year. Using the original dataframe with 2024 registrants, use `mutate()` and lubridate to produce week numbers from the COUNTY_REGISTRATION_DATE (Lubridate has a function that turns a date into a week number. Use that to answer this question). Save that to the original dataframe. Then calculate how many voters are in each week. Describe the distribution of voters groups in a sentence - are the weeks from July 21-August 7 different from the rest, and if so, how?

**A2** The number of new registrations during the weeks of July 21 through Aug. 7 was lower in comparison to the other weeks for the most part. The data could suggest that this period was an outlier because it shows a sharp change in the amount of new voters.

```{r}
md_voters_2024 <- md_voters_2024 |>
mutate(week = week(COUNTY_REGISTRATION_DATE))

md_voters_2024 |>
group_by(week) |>
summarize(total_voters = n()) |>
arrange(week)
```

**Q3** Using lubridate, show which month saw the largest number of registrations. Does the answer help shed more light on Q2?

**A3** This data shows that the largest amount of registrations occured on week 7, with a significantly high amount of registrations. This leads me to wonder how the July 21-August 7 data fits into the registration of the other weeks.

```{r}
md_voters_2024 |>
mutate(month = month(COUNTY_REGISTRATION_DATE)) |>
group_by(month) |>
summarize(total_voters = n()) |>
arrange(desc(total_voters))
```

**Q4** Maryland has, officially, five regions: https://www.visitmaryland.org/article/maryland-regions. Modify the original dataframe again to add a `region` column based on the `COUNTY` value using case_when. Are you going to have to do this for every Maryland county plus Baltimore City? Yes, you are.

Which regions and parties account for the most new voters?

**A4** Democrats in the Capital Region seem to account for the most new voters.

```{r}
md_voters_2024 <- md_voters_2024 |>
mutate(
region = case_when(
COUNTY %in% c("Allegany", "Garrett", "Washington") ~ "Western Maryland",
COUNTY %in% c("Frederick", "Montgomery", "Prince George's") ~ "Capital Region",
COUNTY %in% c("Anne Arundel", "Baltimore", "Baltimore City", "Carroll", "Harford", "Howard") ~ "Central Maryland",
COUNTY %in% c("Calvert", "Charles", "St. Mary's") ~ "Southern Maryland",
COUNTY %in% c("Caroline", "Cecil", "Dorchester", "Kent", "Queen Anne's", "Somerset", "Talbot", "Wicomico", "Worcester") ~ "Eastern Shore"
    )
  )

new_voters <- md_voters_2024 |>
filter(STATE_REGISTRATION_DATE == COUNTY_REGISTRATION_DATE) |>
filter(COUNTY_REGISTRATION_DATE > mdy("07/21/2024"))

new_voters |>
group_by(region, PARTY) |>
summarize(total_new = n()) |>
arrange(desc(total_new))
```


**Q5** Filter md_voters_2024 to only include voters from the Capital region or the Western region. Save the data in a new variable/data set named capital_or_western.

How many voters are in the new data set?

```{r}
capital_or_western <- md_voters_2024 |>
filter(region == "Capital Region" | region == "Western Maryland")

nrow(capital_or_western)

```
Answer: 49052

**Q6** Finally, filter md_voters_2024 to only include voters who meet BOTH of these conditions:

1) they are from the Eastern Shore region or Western region
2) their party affiliation is DEM and their Gender is F

Hint: you can accomplish this with two filters

Save the data in a new variable/data set named final_filter.

How many voters are in the new data set?

```{r}
final_filter <- md_voters_2024 |>
filter(region == "Eastern Shore" | region == "Western Maryland") |>
filter(PARTY == "DEM" & GENDER == "F")

nrow(final_filter)
```

Answer: 1736 are both Democrats and Female


**BONUS QUESTION** What's an additional journalistic question you could ask of this data? Show me the code you'd use to do it below, or at least describe the logic step-by-step in English if you're not sure how to write the code for it yet.


**A** Another question I could ask of this data is the age of the newly registered voters. I think it could be interesting to find out what the average age of newly registered voters is? Are more young people registering to vote?

```{r}

```

